### 2022-09-20 main/chapter_context.tex
add some crap sm notes

The \atlas\ detector did not assemble itself into a preordained form.
It was designed with intent to provide maximally useful and novel data
within real-world cost budgets.
Original designs hoped to construct an \atlas\ detector that could
observe the Higgs boson (which was predicted but yet unseen),
precisely measure known massive objects ($W^\pm$, $Z$, $t$),
and possibly discover new particles such as those hypothesized by
supersymmetric models~\cite{atlas1999design1}.
We now review these three goals with the subsequent news from \atlas'
operations.

\paragraph{Higgs boson discovery.}
Prior to \atlas' construction, the Higgs boson mass (and existence) was
poorly known, but loosely constrained by results like
$m_H \gtrsim 90\,\eV[G]$ from searches at the
Large Electron–Positron Collider (LEP),
$m_H = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
and $m_H \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
atlas1999design2,
ghinculov1998perturb,
lep1999ewk
}.
The Higgs boson couples to an extraordinary variety of fundamental particles,
and \atlas' designers achieved discovery sensitivity across the entire
$100$ to $1000\,\eV[G]$
window by targetting decays including
$H\to b\bar b$,
$H\to \gamma\gamma$,
$H\to ZZ^* \to 4\ell$, and
$H\to ZZ \to \ell\ell\nu\nu$.
This spectrum demands precise measurements of leptons, photons, and
$b$-jets, as well as sensitivity to the missing transverse momentum of
invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays at the very highest
masses.

@@

\paragraph{Precise measurements of known objects.}
\TODO{The top quark's almost-exclusive decay to a $b$ ... $b$ sensitivity}
% top physics: tttt, ttH
% rare ZZy ?
% W mass measurement

>>>>

The \atlas\ detector was not delivered in a preordained form.
It was designed with intent to provide maximally useful and novel data within
real-world cost budgets.
Original designs hoped to construct an \atlas\ detector that could
observe the Higgs boson (which was predicted but yet unseen),
precisely measure known massive objects ($W$, $Z$, and $t$),
and possibly discover new particles such as those hypothesized by
supersymmetric models~\cite{atlas1999design1}.
We now review these three goals with the subsequent news from \atlas'
operations.

\paragraph{Higgs boson discovery.}
Prior to \atlas' construction, the Higgs boson mass (and existence) was
poorly known, but loosely constrained by results like
$m_H \gtrsim 90\,\eV[G]$ from searches at the
Large Electron–Positron Collider (LEP),
$m_H = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
and $m_H \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
atlas1999design2,
ghinculov1998perturb,
lep1999ewk
}.
The Higgs boson couples to an extraordinary variety of fundamental particles.
For its discovery alone, \atlas' designers achieved sensitivity across the
entire $100$ to $1000\,\eV[G]$ window by targetting decays including
$H\to b\bar b$,
$H\to \gamma\gamma$,
$H\to ZZ^* \to 4\ell$, and
$H\to ZZ \to \ell\ell\nu\nu$;
these channels demand precise measurements of leptons, photons, and
$b$-jets, as well as sensitivity to the missing transverse momentum of
invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays at the very highest
masses.

@@

\paragraph{Measurement of known massive objects.}
Other than the Higgs boson, the heavy Standard Model resonances are the
$W$ and $Z$ bosons and the top quark, all of which are produced at
high rates by LHC proton-proton collisions.

The goal of measuring these heavy objects puts more emphasis on determining
the total energy of hadronic jets than their precise hadronic compositions.

\TODO{Although \atlas\ has an active B-physics programme, the stronger particle
identification of the \lhcb\ detector makes it more competitive for hadronic
physics.?}
\TODO{Heavy ions and ALICE?}

\TODO{standard model~\cite{atlas2021summarysm}}

\TODO{top mass~\cite{atlas2022symmarytop, atlas2019topmass}}

\TODO{Top factory. The top quark's almost-exclusive decay as $t\rightarrow Wb$ ... $b$ sensitivity}
% top physics: tttt, ttH
% rare ZZy ?
% W mass measurement

\TODO{$W$ mass~\cite{atlas2018wmass}}
\TODO{`diboson' $WZ$ $W\gamma$}

\TODO{B-physics (supplements dedicated results from \lhcb)}

$\twoljets$ piggyback on the $ZZ \to \ell\ell\nu\nu$ channel
